Bayesian Data Analysis - Rough Draft

### 2023-11-30 Lahman Analysis/Formal Analysis/Draft 1.qmd
Ran the model
Ran the big model and saved the RDS locally becasue GitHub won't let me upload it. I will use it to reference my figures which will mean that only I can view it, so we may have to find some way that I can transport it to Dr. Parson if he wants to take a peek.

#Calculated each players HR proportion "p"
Batting <- Batting %>% 
  mutate(p = HR / AB, .after=AB)
#You can see this for yourself too, but there are 657 players in this dataframe now
```

```{r}
#Choose 500 players at random to consider
#If this doesn't take too long, then we can come back later and change this to possibly consider all players. Or just widen the initial time-frame considered and do 1000 players

@@

```{r}
#Make a dataframe that only keeps the 500 players
Samp_Batting <- Batting %>% 
  filter(playerID %in% N_player)

@@

DISCUSS THE OTHER ARTICLE THAT DR. PARSON SHARED WITH YOU ON SIMILAR/SAME RESEARCH

## Model

@@

@@

$$
\begin{align}
HR_n &\sim Binomial(AB_n,p)\\
logit(p) &= \beta_{0}Player_n+\beta_1cAge_n+\beta_2cAge^2_n+\beta_3cAge^3_n+\beta_4cAge^4_n+\beta_5Park
\end{align}
$$

$$
\begin{align}
\beta_0&\sim Normal(\mu_0, \sigma_0)\\
\beta_1&\sim Normal(\mu_1, \sigma_1)\\
\beta_2&\sim Normal(\mu_2, \sigma_2)\\
\beta_3&\sim Normal(\mu_3, \sigma_3)\\
\beta_4&\sim Normal(\mu_4, \sigma_4)\\
\end{align}
$$

$$
\begin{align}
\mu_0 &\sim Normal(-3.5, 0.1)\\
\sigma_0&\sim Exponential(1)
\end{align}
$$

@@

We recognize that this depends on the hitter which is where the use of a multilevel model becomes useful. Instead of simply using the average probability as our intercept, which is what a dummy variable would do in a GLM, we are able to allow our model choose an intercept which agrees with the data but can adapt depending on the player. Thus, we let $\beta_0$ be normally distributed with mean $\mu_0$ (where $\mu_0$ is normally distributed with mean -3.5 and standard deviation 0.1) and standard deviation $\sigma_0$ (where the parameter $\sigma_0$ itself is exponentially distributed).

@@

>>>>

#Calculated each players HR proportion "p"
Batting <- Batting %>% 
  mutate(p = HR / AB, .after=AB)
#You can see this for yourself too, but there are 657 players in this dataframe now

#Create an index for each player
Batting <- Batting %>% 
  mutate(player_index = as.factor(playerID), .after = playerID) %>% 
  mutate(player_index = as.integer(player_index))

#Create a centered age variable by subtracting 30 from everyone's age

Batting <- Batting %>% 
  mutate(cAge = Age -30, .after = Age)
```

```{r}
#| eval: false
#Choose 500 players at random to consider
#If this doesn't take too long, then we can come back later and change this to possibly consider all players. Or just widen the initial time-frame considered and do 1000 players

@@

```{r}
#| eval: false
#Make a dataframe that only keeps the 500 players
Samp_Batting <- Batting %>% 
  filter(playerID %in% N_player)

@@

# Model

@@

```{r}
#Code to see if HRPolyFit still runs
stan(
  file = "HRPolyFit.stan",
  data = data.poly.fit,
  iter = 100
)
```

@@

$$
\begin{align}
HR_{ni} &\sim Binomial(AB_{ni},p_{ni})\\
logit(\frac{p_{ni}}{1-p_{ni}}) &= \alpha_{n}+\beta_h+\eta_h+\delta_p+\xi_i
\end{align}
$$

$$
\begin{align}
\alpha_n&\sim Normal(\mu_0, \sigma_0)\\
\beta_h&\sim Normal(\mu_1, \sigma_1)\\
\eta_h&\sim Normal(\mu_2, \sigma_2)\\
\delta_p&\sim Normal(\mu_5, \sigma_5)\\
\xi_i&\sim Normal(\mu_6, \sigma_6)\\
\end{align}
$$

$$
\begin{align}
\mu_0 &\sim Normal(-3.5, 0.1)\\
\sigma_0&\sim Exponential(1)\\
\\
\mu_1&\sim Normal(0,0.01)\\
\sigma_1&\sim Exponential(10)\\
\\
\mu_2&\sim Normal(0,0.01)\\
\sigma_2&\sim Exponential(100)\\
\\
\mu_5&\sim Normal(0,0.01)\\
\sigma_5&\sim Exponential(10)\\
\\
\mu_6&\sim Normal(0,0.25)\\
\sigma_6&\sim Exponential(10)
\end{align}
$$

@@

We recognize that this depends on the hitter which is where the use of a multilevel model becomes useful. Instead of simply using the average probability as our intercept, which is what a dummy variable would do in a GLM, we are able to allow our model choose an intercept which agrees with the data but can adapt depending on the player. Thus, we let $\alpha$ be normally distributed with mean $\mu_0$ (where $\mu_0$ is normally distributed with mean -3.5 and standard deviation 0.1) and standard deviation $\sigma_0$ (where the parameter $\sigma_0$ itself is exponentially distributed).

$\beta_h$ the effect that a player being age $h$ has on HR probability. Similarly $\eta_h$ is the effect that a player's age squared has on the probability of a HR in an AB. This squared term captures the non-linearity of HR as player's age.

$\delta_p$ is the effect that a player's home ballpark $p$ has on his underlying probability of hitting a home run.

$\xi_i$ represents the change in the underlying probability of player hitting a HR in the year $i$

@@

# Results

```{r}
#| eval: false

#Since we aren't fitting anything when we render, we don't need this chunk to take up any extra time calculating usless vectors
data.poly.fit <- list(
  N = nrow(Batting),
  n_player = n_distinct(Batting$player_index),
  n_park = n_distinct(Batting$park_index),
  n_year = n_distinct(Batting$yearID),
  AB = Batting %>% pull(AB),
  HR = Batting %>% pull(HR),
  Player = Batting %>% pull(player_index),
  cAge = Batting %>% pull(cAge),
  cAgeSq = (Batting %>% pull(cAge))**2,
  Park = Batting %>% pull(park_index),
  Year = Batting %>% pull(yearID) -1972
)
```

```{r}
#| eval: false

#We don't want the program, which takes my desktop 30-45 minutes to render, to run every time I need to render the docuement
fit.poly <- stan(
  file = "HRPolyFit.stan",
  data = data.poly.fit,
  iter = 1e4
)
```

```{r}
#| eval: false

#This saves the fit as and RDS so that I can reference it later when rendering
saveRDS(fit.poly, "PolyFit.rds")
```

```{r}
plot(fit.poly, pars="xi")
```
